---
title: "cartogram: Create Cartograms with R"
output:
  github_document:
    fig_width: 5
    fig_height: 5
---

[![Build Status](https://travis-ci.org/sjewo/cartogram.svg?branch=master)](https://travis-ci.org/sjewo/cartogram)
[![CRAN Downloads](http://cranlogs.r-pkg.org/badges/cartogram)](https://cran.r-project.org/package=cartogram)

```{r, echo=F}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "tools/readme/README-"
)
```

Construct a continuous area cartogram by a rubber sheet distortion algorithm (Dougenik et al. 1985) or non-contiguous Area Cartograms (Olson 1976) in R.

## Installation

You can install the **cartogram** package from CRAN as follows:

```{r, eval=FALSE}
install.packages("cartogram")
```

To upgrade to the latest development version of `cartogram`, install the package `devtools` and run the following command:

```{r, eval=FALSE}
devtools::install_github("sjewo/cartogram")
```

## NEWS

* [0.1.0] Non-Overlapping Circles Cartogram (sf and sp)
* [0.0.3] sf support added
* [0.0.2] Non-contiguous Area Cartogram
* [0.0.2] Prepare data with missing or extreme values before cartogram calculation for faster convergence
* [0.0.1] Initial Release

## Examples

### Continuous Area Cartogram

```{r carto}
library(cartogram)
library(tmap)
library(maptools)

data(wrld_simpl)

afr <- wrld_simpl[wrld_simpl$REGION==2,]
afr <- spTransform(afr, CRS("+init=epsg:3395"))

# construct cartogram
afrc <- cartogram_cont(afr, "POP2005", itermax=5)

# plot it
tm_shape(afrc) + tm_fill("POP2005", style="jenks") + 
  tm_borders() + tm_layout(frame=F)
```


### Non-contiguous Area Cartogram
Many thanks to @rCarto and @neocarto for contributing the code!

```{r nc}
library(cartogram)
library(tmap)
library(maptools)

data(wrld_simpl)

afr <- wrld_simpl[wrld_simpl$REGION==2,]
afr <- spTransform(afr, CRS("+init=epsg:3395"))

# construct cartogram
afrnc <- cartogram_ncont(afr, "POP2005")

# plot it
tm_shape(afr) + tm_borders() + 
  tm_shape(afrnc) + tm_fill("POP2005", style="jenks") + 
  tm_borders() + tm_layout(frame=F)
```

### Non-Overlapping Circles Cartogram
Many thanks to @rCarto for contributing the code!

```{r noc}
library(cartogram)
library(tmap)
library(maptools)

data(wrld_simpl)

afr <- wrld_simpl[wrld_simpl$REGION==2,]
afr <- spTransform(afr, CRS("+init=epsg:3395"))

# construct cartogram
afrnoc <- cartogram_noc(afr, "POP2005")

# plot it
tm_shape(afr) + tm_borders() + 
  tm_shape(afrnoc) + tm_fill("POP2005", style="jenks") + 
  tm_borders() + tm_layout(frame=F)
```


## sf support

Thanks to @Nowosad for speeding things up!

```{r sfsupport, fig.width=7}
library(sf)
library(cartogram)
library(maptools)

data(wrld_simpl)

afr <- wrld_simpl[wrld_simpl$REGION==2,]
afr <- spTransform(afr, CRS("+init=epsg:3395"))

afr_sf = st_as_sf(afr)

# Display plots in two columns
par(mfrow=c(1,2), mai=c(0,0,0,0))

# Continuous Area Cartogram
afr_sf_carto <- cartogram_cont(afr_sf, "POP2005", 3)
plot(st_geometry(afr_sf_carto))

# Plot Non-contiguous Area Cartogram
afr_sf_nc <- cartogram_ncont(afr_sf, "POP2005")
plot(st_geometry(afr_sf))
plot(st_geometry(afr_sf_nc), add = TRUE, col = 'red')
```



## References

* Dorling, D. (1996). Area Cartograms: Their Use and Creation. In Concepts and Techniques in Modern Geography (CATMOG), 59.
* Dougenik, J. A., Chrisman, N. R., & Niemeyer, D. R. (1985). An Algorithm To Construct Continuous Area Cartograms. In The Professional Geographer, 37(1), 75-81.
* Olson, J. M. (1976), Noncontiguous Area Cartograms. The Professional Geographer, 28: 371–380. doi:10.1111/j.0033-0124.1976.00371.x
