---
title: "cartogram: Create Cartograms with R"
output:
  github_document:
    fig_width: 4
    fig_height: 3.5
---

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/cartogram)](https://cran.r-project.org/package=cartogram)
[![R-CMD-check](https://github.com/sjewo/cartogram/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/sjewo/cartogram/actions/workflows/R-CMD-check.yaml)
[![CRAN Downloads](https://cranlogs.r-pkg.org/badges/cartogram)](https://cran.r-project.org/package=cartogram)
<!-- badges: end -->

```{r, echo=F}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

Construct a continuous area cartogram by a rubber sheet distortion algorithm (Dougenik et al. 1985), non-contiguous Area Cartograms (Olson 1976), and non-overlapping Circles Cartogram (Dorling el al. 1996) in R.

## Installation

You can install the **cartogram** package from CRAN as follows:

```{r, eval=FALSE}
install.packages("cartogram")
```

To upgrade to the latest development version of `cartogram`, install the package `remotes` and run the following command:

```{r, eval=FALSE}
remotes::install_github("sjewo/cartogram")
```

## Examples

### Continuous Area Cartogram

```{r cont, fig.asp = 1.2}
library(cartogram)
library(sf)
library(tmap)

data("World")

# keep only the african continent
afr <- World[World$continent == "Africa", ]

# project the map
afr <- st_transform(afr, 3395)

# construct cartogram
afr_cont <- cartogram_cont(afr, "pop_est", itermax = 5)

# plot it
tm_shape(afr_cont) + tm_polygons("pop_est", style = "jenks") +
  tm_layout(frame = FALSE, legend.position = c("left", "bottom"))
```

### Non-contiguous Area Cartogram

Many thanks to @rCarto and @neocarto for contributing the code!

```{r ncont, fig.asp = 1.2}
# construct cartogram
afr_ncont <- cartogram_ncont(afr, "pop_est")

# plot it
tm_shape(afr) + tm_borders() +
  tm_shape(afr_ncont) + tm_polygons("pop_est", style = "jenks") +
  tm_layout(frame = FALSE, legend.position = c("left", "bottom"))
```

### Non-Overlapping Circles Cartogram

Many thanks to @rCarto for contributing the code!

```{r dorling, fig.asp = 1.2}
# construct cartogram
afr_dorling <- cartogram_dorling(afr, "pop_est")

# plot it
tm_shape(afr) + tm_borders() +
  tm_shape(afr_dorling) + tm_polygons("pop_est", style = "jenks") +
  tm_layout(frame = FALSE, legend.position = c("left", "bottom"))
```

## Use multiple CPU cores

Many thanks to @e-kotov for contributing the code!

```{r parallel, fig.asp = 1.2}
library(cartogram)
library(sf)
library(tmap)

library(future)
library(future.apply)
library(parallelly)

data("World")

# keep only the african continent
afr <- World[World$continent == "Africa", ]

# project the map
afr <- st_transform(afr, 3395)

# Create cartogram using 2 CPU cores on local machine
afr_cont <- cartogram_cont(afr, weight = "pop_est", itermax = 5, n_cpu = 2)

# plot it
tm_shape(afr_cont) + tm_polygons("pop_est", style = "jenks") +
  tm_layout(frame = FALSE, legend.position = c("left", "bottom"))
```

## References

* Dorling, D. (1996). Area Cartograms: Their Use and Creation. In Concepts and Techniques in Modern Geography (CATMOG), 59.
* Dougenik, J. A., Chrisman, N. R., & Niemeyer, D. R. (1985). An Algorithm To Construct Continuous Area Cartograms. In The Professional Geographer, 37(1), 75-81.
* Olson, J. M. (1976), Noncontiguous Area Cartograms. The Professional Geographer, 28: 371–380. [doi:10.1111/j.0033-0124.1976.00371.x](https://doi.org/10.1111/j.0033-0124.1976.00371.x)
